---
Title: Propensity Score Matching and study of Writing Lab Impact
author: "Meng Deng"
output: pdf_document
---

#Propensity Score Matching and study of Writing Lab Impact
##Meng Deng 
In this chapter, we perform propensity score matching to form a quasi-experiment to investigate on the writing lab impact on students' academic performance. 

Matching is performed using the MatchIt package. To be specific, we followed the following steps. 
  
  1. Perform pre-analysis using non-matched data. We first exam the effect by writing lab status. The means difference are supposed to significant by writing lab status.
    1. Investigate means difference on outcome variables (gpa by wl_indicator).
    2. Investigate means difference on pre-treatment covariates (i.e. gender by wl_indicator).
  
  2. Estimate the propensity score by running a logit model on both int'l and domestic datasets. We come up with the propensity score for each student. 
    1. Domestic: $log\left(\frac{visit WL}{not visit WL} \right) = \beta_0$ + $first_term_gpa$ + $gender$ + $avg_first_view$ + $ethnicity$ + $profile_reading_pct$
    2. International: $log\left(\frac{visit WL}{not visit WL} \right) = \beta_0$ + $first_term_gpa$ + $gender$ + $toefl$ + $profile_reading_pct$

  3. Pair observations that have very similar propensity scores, but that differ in their writing lab status. 

  4. Assess covariate balance in the matched datasets. 
    1. Perform visual inspection by plotting the mean of each covariate against the estimated propensity score, separately by treatment status. The treatment and control groups are supposed to have identical means of each covariate at each value of the propensity score for successful matching.
    2. t-tests (or chi-squire for categorical variables) of difference-in-means. For well matching, we should not be able to reject the null hypothesis of no mean difference for each covariate. 

  5. Investigate treatment effects using matched datasets.
    1. Two-way ANOVA models. 
    2. Logistic regression model.
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r echo=FALSE}
library(psych)
library(BSDA)
library(glmnet)
library(MatchIt)
library(ggplot2)
library(gridExtra)
library(nnet)
library(effects)
library(lsmeans)
library(plyr)
library(ggplot2)
library(multcompView)


setwd('~/Dropbox/STAT 526/526 Group Project/Processed Data')

#-------------------Domestic students--------------------------------- 
imputed_dom <- read.csv('Imputed_domestic_new.csv', sep=',')

imputed_dom2 <- subset(imputed_dom, select=-c(visit_times, profile_academic_period, PROFILE_RESIDENCE))


#factorize and ordinal factors. 
imputed_dom2$wla_indicator <- as.factor(imputed_dom2$wla_indicator)

imputed_dom2$profile_read_pct <- ordered(imputed_dom2$profile_read_pct)
imputed_dom2$profile_read_pct <- ordered(imputed_dom2$profile_read_pct, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))

imputed_dom2$profile_write_pct <- ordered(imputed_dom2$profile_write_pct)
imputed_dom2$profile_write_pct <- ordered(imputed_dom2$profile_write_pct, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))

imputed_dom2$profile_math_pct <- ordered(imputed_dom2$profile_math_pct)
imputed_dom2$profile_math_pct <- ordered(imputed_dom2$profile_math_pct, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))


imputed_dom2$at_risk_ind_term <- as.factor(imputed_dom2$at_risk_ind_term)


#----------------------International students--------------------------------
imputed_intl <- read.csv('Imputed_international_new.csv', sep=',')

imputed_intl2 <- subset(imputed_intl, select=-c(visit_times, profile_academic_period, profile_residence))


#factorize and ordinal factors. 
imputed_intl2$wla_indicator <- as.factor(imputed_intl2$wla_indicator)

imputed_intl2 <- subset(imputed_intl2, ttd_class_label != 'norm')

imputed_intl2$profile_read_pct <- ordered(imputed_intl2$profile_read_pct)
imputed_intl2$profile_read_pct <- ordered(imputed_intl2$profile_read_pct, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))

imputed_intl2$profile_write_pct <- ordered(imputed_intl2$profile_write_pct)
imputed_intl2$profile_write_pct <- ordered(imputed_intl2$profile_write_pct, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))

imputed_intl2$profile_math_pct <- ordered(imputed_intl2$profile_math_pct)
imputed_intl2$profile_math_pct <- ordered(imputed_intl2$profile_math_pct, levels = c(">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))

imputed_intl2$toefl_score <- ordered(imputed_intl2$toefl_score)
imputed_intl2$toefl_score <- ordered(imputed_intl2$toefl_score, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=40%", ">=55%", ">=70%", ">=85%", ">=96%"))

imputed_intl2$at_risk_ind_term <- as.factor(imputed_intl2$at_risk_ind_term)

```

##1. Analysis of pre-matching data. We first exam the effect by writing lab status. 

We first exam the outcome variables (gpa and ttd_class_label) by wl status. 
```{r echo=FALSE}

#----------------------------------Domestic students--------------------------
#Any difference on gpa by wla_indicator? Yes.  
describeBy(imputed_dom2$gpa, imputed_dom2$wla_indicator)
tsum.test(mean.x=2.98, s.x=.67, n.x=100872,
          mean.y=3.16, s.y=.51, n.y=11469)

#Any difference on ttd status by wla_indicator? Yes.
with(imputed_dom2, table(wla_indicator, ttd_class_label))
with(imputed_dom2, chisq.test(ttd_class_label, wla_indicator))

#----------------------------------International students----------------------
#Any difference on gpa by wla_indicator? Yes.  
describeBy(imputed_intl2$gpa, imputed_intl2$wla_indicator)
tsum.test(mean.x=3, s.x=.75, n.x=13362,
          mean.y=3.29, s.y=.49, n.y=6407)

#Any difference on ttd status by wla_indicator? Yes.
with(imputed_intl2, table(wla_indicator, ttd_class_label))
with(imputed_intl2, chisq.test(ttd_class_label, wla_indicator))
```

We now want to exam the covriates by wl status, instead of investigating on all the variables, we want to run the lasso process to select variables before testing their effects on wl status.
Variables entered the selection process for domestic students:
 * gender
 * underrepresented_minority_ind
 * reporting_ethnicity
 * hs_core_gpa
 * avg_scholarship_amt_per_term
 * scholarship_award_term_count
 * PROFILE_RESIDENCE
 * profile_read_pct
 * profile_write_pct
 * profile_math_pct
 * avg_z_first_view_ts
 * first_term_gpa
 * avg_not_in_class_hours_per_devic
 * college_decode
 
We now want to exam the covriates by wl status, instead of investigating on all the variables, we want to run the lasso process to select variables before testing their effects on wl status.
Variables entered the selection process for international students:
 * gender
 * hs_core_gpa
 * profile_read_pct
 * profile_write_pct
 * profile_math_pact
 * toefl_score
 * avg_not_in_class_hours_per_device
 * college_decode
 * first_term_gpa
 
```{r echo=FALSE}
#--------------------------------Domestic students-----------------------------------
#create a matrix which creating dummies for all factor independent variables.
dummies <- model.matrix(wla_indicator~gender + underrepresented_minority_ind + reporting_ethnicity +  profile_residence + profile_read_pct + profile_write_pct + profile_math_pct + college_decode + at_risk_ind_term, imputed_dom2)[,-1]

#combine the dummy with the original df.
covariateMatrix <- as.matrix(data.frame(imputed_dom2$hs_core_gpa, imputed_dom2$avg_scholarship_amt_per_term, imputed_dom2$scholarship_award_term_count, imputed_dom2$avg_z_first_view_ts, imputed_dom2$first_term_gpa, imputed_dom2$avg_not_in_class_hours_per_devic, dummies))


fit_lasso <- glmnet(covariateMatrix, y=imputed_dom2$wla_indicator, alpha=1, family="binomial")

#plot variable coefficients vs. shrinkage parameter lambda.
plot(fit_lasso, xvar="lambda")
grid()

fit_lasso
coef(fit_lasso)[, 10]

#---------------------------------International students------------------------------
#create a matrix which creating dummies for all factor independent variables.
dummies <- model.matrix(wla_indicator~gender + profile_read_pct + profile_write_pct + profile_math_pct + toefl_score + college_decode + at_risk_ind_term, imputed_intl2)[,-1]

#combine the dummy with the original df.
covariateMatrix <- as.matrix(data.frame(imputed_intl2$hs_core_gpa, imputed_intl2$avg_z_first_view_ts, imputed_intl2$avg_not_in_class_hours_per_devic, imputed_intl2$first_term_gpa, dummies))


fit_lasso <- glmnet(covariateMatrix, y=imputed_intl2$wla_indicator, alpha=1, family="binomial")

#plot variable coefficients vs. shrinkage parameter lambda.
plot(fit_lasso, xvar="lambda")
grid()

fit_lasso
coef(fit_lasso)[, 10]

```

As a result, the lasso procedure selected the following variables for the matching process. We then investigate these four covariates by treatment effect (wla_indicator). We found significant effect by wla_indicator in all of the covariates.
Variable after the selection for domestic students:
* first_term_gpa
* avg_z_first_view_ts
* gender
* reporting_ethnicity
* profile_read_pct
* profile_write_pct

Variable after the selection for international students:
* first_term_gpa
* gender
* profile_read_pct
* toefl_score

```{r echo=FALSE}
#-------------------------Domestic students-----------------------------------
#Use t-test to test on first_term_gpa by wla_indicator.
describeBy(imputed_dom2$first_term_gpa, imputed_dom2$wla_indicator)

tsum.test(mean.x=2.84, s.x=0.84, n.x=100872,
          mean.y=2.92, s.y=0.78, n.y=11469)

#Use t-test to test on first_term_gpa by wla_indicator.
describeBy(imputed_dom2$avg_z_first_view_ts, imputed_dom2$wla_indicator)

tsum.test(mean.x=0.05, s.x=0.52, n.x=100872,
          mean.y=-0.03, s.y=0.5, n.y=11469)

#Use chi-square to test gender by wla_indicator.
with(imputed_dom2, table(wla_indicator, gender))

with(imputed_dom2, chisq.test(gender, wla_indicator))

#Use chi-square to test reporting_ethnicity by wla_indicator.
with(imputed_dom2, table(wla_indicator, reporting_ethnicity))

with(imputed_dom2, chisq.test(reporting_ethnicity, wla_indicator))

#Use chi-square to test profile_read_pct by wla_indicator.
with(imputed_dom2, table(wla_indicator, profile_read_pct))

with(imputed_dom2, chisq.test(wla_indicator, profile_read_pct))

#Use chi-square to test profile_write_pct by wla_indicator.
with(imputed_dom2, table(wla_indicator, profile_write_pct))

with(imputed_dom2, chisq.test(wla_indicator, profile_write_pct))

#------------------------------international students---------------------------
#Use t-test to test on at_risk_ind_term by wla_indicator.
describeBy(imputed_intl2$first_term_gpa, imputed_intl2$wla_indicator)

tsum.test(mean.x=2.94, s.x=0.95, n.x=13362,
          mean.y=3.26, s.y=0.75, n.y=6407)

#Use chi-square to test gender by wla_indicator.
with(imputed_intl2, table(wla_indicator, gender))

with(imputed_intl2, chisq.test(gender, wla_indicator))

#Use chi-square to test profile_read_pct by wla_indicator.
with(imputed_intl2, table(wla_indicator, profile_read_pct))

with(imputed_intl2, chisq.test(wla_indicator, profile_read_pct))

#Use chi-square to test toefl_score by wla_indicator.
with(imputed_intl2, table(wla_indicator, toefl_score))

with(imputed_intl2, chisq.test(wla_indicator, toefl_score))
```

##2 Matching process.

###2.1. Propensity score estimation. 

We estimate propensity scores by running a logit model where the outcome variable is a binary variable indicating treatment status. We include the covariates from the lasso model selection results.

```{r echo=FALSE}
#-----------------------------------domestic students----------------------------------
imputed_dom2_ps <- glm(wla_indicator ~ first_term_gpa + gender + avg_z_first_view_ts + reporting_ethnicity + profile_read_pct + profile_write_pct,
            family = 'binomial', imputed_dom2)
#summary(imputed_dom2_ps)

#-----------------------------------international students-----------------------------
imputed_intl2_ps <- glm(wla_indicator ~ first_term_gpa + gender + profile_read_pct + toefl_score, family = 'binomial', imputed_intl2)

```

Now calculate the propensity score for each student. Proposentiy score is the student's predicted probability of visiting the writing lab, given the estimates from the logit model.

```{r echo=FALSE}
#-----------------------------------domestic students----------------------------------
imputed_dom2_df <- data.frame(pr_score = predict(imputed_dom2_ps, type = "response"), wla_indicator = imputed_dom2_ps$model$wla_indicator)

#-----------------------------------international students-----------------------------
imputed_intl2_df <- data.frame(pr_score = predict(imputed_intl2_ps, type = "response"), wla_indicator = imputed_intl2_ps$model$wla_indicator)
```

We can plot histograms of the estimated propensity scores by treatment status to exam the region of common support.

```{r echo=FALSE}
#-----------------------------------domestic students----------------------------------
labs <- paste("Actual writing lab visiting status:", c("Writing lab visitors", "Not writing lab users"))
imputed_dom2_df$wla_indicator = ifelse(imputed_dom2_df$wla_indicator == 1, labs[1], labs[2])

ggplot(data=imputed_dom2_df, aes(x=pr_score, group=wla_indicator)) + geom_histogram(color = "white") + facet_wrap(~wla_indicator) + xlab("Probability of visiting writing lab") + theme_bw()

#-----------------------------------international students-----------------------------
labs <- paste("Actual writing lab visiting status:", c("Not writing lab users", "Writing lab visitors"))
imputed_intl2_df$wla_indicator = ifelse(imputed_intl2_df$wla_indicator == 0, labs[1], labs[2])

ggplot(data=imputed_intl2_df, aes(x=pr_score, group=wla_indicator)) + geom_histogram(color = "white") + facet_wrap(~wla_indicator) + xlab("Probability of visiting writing lab") + theme_bw()
```

###2.2. Propensity score matching.

```{r echo=FALSE}
#-----------------------------------domestic students----------------------------------
imputed_dom2_match <- matchit(wla_indicator ~ first_term_gpa + gender + avg_z_first_view_ts + reporting_ethnicity + profile_read_pct + profile_write_pct, method = "nearest", data = imputed_dom2)
summary(imputed_dom2_match)
plot(imputed_dom2_match)

#Produce the matched dataset
imputed_dom2_match_after <- match.data(imputed_dom2_match)

imputed_dom2_match_after$wla_indicator <- as.factor(imputed_dom2_match_after$wla_indicator)
imputed_dom2_match_after$wla_indicator <- factor(imputed_dom2_match_after$wla_indicator,
levels = c(0,1),
labels = c("not WL users", "WL users"))


imputed_dom2_match_after$profile_read_pct <- ordered(imputed_dom2_match_after$profile_read_pct)
imputed_dom2_match_after$profile_read_pct <- ordered(imputed_dom2_match_after$profile_read_pct, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))

imputed_dom2_match_after$profile_write_pct <- ordered(imputed_dom2_match_after$profile_write_pct)
imputed_dom2_match_after$profile_write_pct <- ordered(imputed_dom2_match_after$profile_write_pct, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))

imputed_dom2_match_after$profile_math_pct <- ordered(imputed_dom2_match_after$profile_math_pct)
imputed_dom2_match_after$profile_math_pct <- ordered(imputed_dom2_match_after$profile_math_pct, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))



imputed_dom2_match_after$at_risk_ind_term <- as.factor(imputed_dom2_match_after$at_risk_ind_term)
imputed_dom2_match_after$at_risk_ind_term <- factor(imputed_dom2_match_after$at_risk_ind_term,
levels = c(0,1),
labels = c("not at risk", "at risk"))

#-----------------------------------international students-----------------------------
imputed_intl2_match <- matchit(wla_indicator ~ first_term_gpa + gender + profile_read_pct + toefl_score, method = "nearest", data = imputed_intl2)
#summary(imputed_intl2_match)
plot(imputed_intl2_match)


#Produce the matched dataset
imputed_intl2_match_after <- match.data(imputed_intl2_match)

imputed_intl2_match_after$wla_indicator <- as.factor(imputed_intl2_match_after$wla_indicator)
imputed_intl2_match_after$wla_indicator <- factor(imputed_intl2_match_after$wla_indicator,
levels = c(0,1),
labels = c("not WL users", "WL users"))

imputed_intl2_match_after$profile_read_pct <- ordered(imputed_intl2_match_after$profile_read_pct)
imputed_intl2_match_after$profile_read_pct <- ordered(imputed_intl2_match_after$profile_read_pct, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))

imputed_intl2_match_after$profile_write_pct <- ordered(imputed_intl2_match_after$profile_write_pct)
imputed_intl2_match_after$profile_write_pct <- ordered(imputed_intl2_match_after$profile_write_pct, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))

imputed_intl2_match_after$profile_math_pct <- ordered(imputed_intl2_match_after$profile_math_pct)
imputed_intl2_match_after$profile_math_pct <- ordered(imputed_intl2_match_after$profile_math_pct, levels = c(">=10%", ">=25%", ">=50%", ">=75%", ">=90%", ">=95%", ">=99%"))

imputed_intl2_match_after$toefl_score <- ordered(imputed_intl2_match_after$toefl_score)
imputed_intl2_match_after$toefl_score <- ordered(imputed_intl2_match_after$toefl_score, levels = c("<5%", ">=5%", ">=10%", ">=25%", ">=40%", ">=55%", ">=70%", ">=85%", ">=96%"))

imputed_intl2_match_after$at_risk_ind_term <- as.factor(imputed_intl2_match_after$at_risk_ind_term)
imputed_intl2_match_after$at_risk_ind_term <- factor(imputed_intl2_match_after$at_risk_ind_term,
levels = c(0,1),
labels = c("not at risk", "at risk"))
```

##3. Covariates balance assessment after matching. 

###3.1. Now plot the mean of each covariate against the estimated propensity score, separately by treatment status. If matching is done well, the treatment and control groups will have (near) identical means of each covariate at each value of the propensity score.
```{r echo=FALSE}
#-----------------------------------domestic students----------------------------------
ggplot(imputed_dom2_match_after, aes(x = imputed_dom2_match_after$distance, y = imputed_dom2_match_after$first_term_gpa, color = wla_indicator)) +
      geom_point(alpha = 0.2, size = 1.3) +
      geom_smooth(method = "loess", se = F) +
      xlab("Propensity score") +
      ylab("first_term_gpa")

ggplot(imputed_dom2_match_after, aes(x = imputed_dom2_match_after$distance, y = imputed_dom2_match_after$gender, color = wla_indicator)) +
      geom_point(alpha = 0.2, size = 1.3) +
      geom_smooth(method = "loess", se = F) +
      xlab("Propensity score") +
      ylab("Gender")

ggplot(imputed_dom2_match_after, aes(x = imputed_dom2_match_after$distance, y = imputed_dom2_match_after$avg_z_first_view_ts, color = wla_indicator)) +
      geom_point(alpha = 0.2, size = 1.3) +
      geom_smooth(method = "loess", se = F) +
      xlab("Propensity score") +
      ylab("avg_z_first_view_ts")

ggplot(imputed_dom2_match_after, aes(x = imputed_dom2_match_after$distance, y = imputed_dom2_match_after$reporting_ethnicity, color = wla_indicator)) +
      geom_point(alpha = 0.2, size = 1.3) +
      geom_smooth(method = "loess", se = F) +
      xlab("Propensity score") +
      ylab("reporting_ethnicity")

ggplot(imputed_dom2_match_after, aes(x = imputed_dom2_match_after$distance, y = imputed_dom2_match_after$profile_read_pct, color = wla_indicator)) +
      geom_point(alpha = 0.2, size = 1.3) +
      geom_smooth(method = "loess", se = F) +
      xlab("Propensity score") +
      ylab("profile_read_pct")

ggplot(imputed_dom2_match_after, aes(x = imputed_dom2_match_after$distance, y = imputed_dom2_match_after$profile_write_pct, color = wla_indicator)) +
      geom_point(alpha = 0.2, size = 1.3) +
      geom_smooth(method = "loess", se = F) +
      xlab("Propensity score") +
      ylab("profile_write_pct")

#-----------------------------------international students-----------------------------
ggplot(imputed_intl2_match_after, aes(x = imputed_intl2_match_after$distance, y = imputed_intl2_match_after$first_term_gpa, color = wla_indicator)) +
      geom_point(alpha = 0.2, size = 1.3) +
      geom_smooth(method = "loess", se = F) +
      xlab("Propensity score") +
      ylab("first_term_gpa")

ggplot(imputed_intl2_match_after, aes(x = imputed_intl2_match_after$distance, y = imputed_intl2_match_after$gender, color = wla_indicator)) +
      geom_point(alpha = 0.2, size = 1.3) +
      geom_smooth(method = "loess", se = F) +
      xlab("Propensity score") +
      ylab("Gender")

ggplot(imputed_intl2_match_after, aes(x = imputed_intl2_match_after$distance, y = imputed_intl2_match_after$profile_read_pct, color = wla_indicator)) +
      geom_point(alpha = 0.2, size = 1.3) +
      geom_smooth(method = "loess", se = F) +
      xlab("Propensity score") +
      ylab("profile_read_pct")

ggplot(imputed_intl2_match_after, aes(x = imputed_intl2_match_after$distance, y = imputed_intl2_match_after$toefl_score, color = wla_indicator)) +
      geom_point(alpha = 0.2, size = 1.3) +
      geom_smooth(method = "loess", se = F) +
      xlab("Propensity score") +
      ylab("toefl_score")
```

###3.2. Difference-in-means.
The means suggests a high degree of balance on the covariates. 
```{r echo=FALSE}
#-----------------------------------domestic students----------------------------------
aggregate(cbind(imputed_dom2_match_after$gender, imputed_dom2_match_after$reporting_ethnicity,
imputed_dom2_match_after$avg_z_first_view_ts, imputed_dom2_match_after$profile_read_pct, imputed_dom2_match_after$profile_write_pct, imputed_dom2_match_after$first_term_gpa), list(imputed_dom2_match_after$wla_indicator), mean)

#-----------------------------------international students-----------------------------
aggregate(cbind(imputed_intl2_match_after$gender, imputed_intl2_match_after$profile_read_pct, imputed_intl2_match_after$toefl_score, imputed_intl2_match_after$first_term_gpa), list(imputed_intl2_match_after$wla_indicator), mean)
```

Conduct t-test for first_term_gpa and chi-square tests for the rests. Ideally, we should not be able to reject the null hypothesis of no mean difference for each covariate.

```{r echo=FALSE}
#-----------------------------------domestic students----------------------------------
#First_term_gpa no longer siginificant. p = 0.06.
describeBy(imputed_dom2_match_after$first_term_gpa, imputed_dom2_match_after$wla_indicator)


tsum.test(mean.x=2.9, s.x=.8, n.x=11469,
          mean.y=2.92, s.y=.78, n.y=11469)

#first view no longer significant. 
describeBy(imputed_dom2_match_after$avg_z_first_view_ts, imputed_dom2_match_after$wla_indicator)


tsum.test(mean.x=-0.02, s.x=.51, n.x=11469,
          mean.y=-0.03, s.y=.5, n.y=11469)

#Gender no longer significant by wl status. 
with(imputed_dom2_match_after, table(wla_indicator, gender))

with(imputed_dom2_match_after, chisq.test(gender, wla_indicator))

#reporting_ethnicity no longer significant. 
with(imputed_dom2_match_after, table(wla_indicator, reporting_ethnicity))

with(imputed_dom2_match_after, chisq.test(reporting_ethnicity, wla_indicator, simulate.p.value = TRUE))

#profile_read_pct no longer significant by wl status. 
with(imputed_dom2_match_after, table(wla_indicator, profile_read_pct))

with(imputed_dom2_match_after, chisq.test(wla_indicator, profile_read_pct))

#profile_write_pct no longer significant by wl status. 
with(imputed_dom2_match_after, table(wla_indicator, profile_write_pct))

with(imputed_dom2_match_after, chisq.test(wla_indicator, profile_write_pct))

#-----------------------------------international students-----------------------------
#First_term_gpa no longer different by wl status.
describeBy(imputed_intl2_match_after$first_term_gpa, imputed_intl2_match_after$wla_indicator)

tsum.test(mean.x=3.26, s.x=.74, n.x=6407,
          mean.y=3.26, s.y=.75, n.y=6407)

#Gender still seem to be significant by wl status (p=0.02), but the the different effect was largely reduced.
with(imputed_intl2_match_after, table(wla_indicator, gender))

with(imputed_intl2_match_after, chisq.test(gender, wla_indicator))

#profile_read_pct no longer significant by wl status. 
with(imputed_intl2_match_after, table(wla_indicator, profile_read_pct))

with(imputed_intl2_match_after, chisq.test(wla_indicator, profile_read_pct))

#toefl_score still marginally significant by wl status (p=0.05), but the different effect was largely reduced.
with(imputed_intl2_match_after, table(wla_indicator, toefl_score))

with(imputed_intl2_match_after, chisq.test(wla_indicator, toefl_score))

```

##4. Estimate WLA_indicator effects. 

###4.1 Dependent variable: GPA
gpa = wla_indicator*at_risk_ind_term
```{r echo=FALSE}
#-----------------------------------domestic students----------------------------------
imputed_dom2_match_after.lm <- lm(gpa ~ wla_indicator * at_risk_ind_term, data=imputed_dom2_match_after)
anova(imputed_dom2_match_after.lm)

#both the main effects and interaction effect are highly significant. 
coefficients(imputed_dom2_match_after.lm)
confint(imputed_dom2_match_after.lm)

#plots: main effect and interaction effect. 
#main effect.

apatheme=theme_bw()+
    theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        axis.line=element_line(),
        text=element_text(family='Times'))
#lsmeans.
leastsquare = lsmeans(imputed_dom2_match_after.lm, 
                      pairwise ~ wla_indicator:at_risk_ind_term,
                      adjust="tukey")
cld(leastsquare, 
    alpha=.05,  
    Letters=letters)

toothInt <- ddply(imputed_dom2_match_after,.(wla_indicator, at_risk_ind_term),summarise, val = mean(gpa))

#plots: main effect and interaction effect. 
#main effect.
ggplot(imputed_dom2_match_after, aes(x = at_risk_ind_term, y = gpa, color=wla_indicator)) + geom_boxplot() + scale_x_discrete() + xlab("at_risk_ind_term") + ylab("gpa") + apatheme 


ggplot(imputed_dom2_match_after.lm, aes(x = wla_indicator, y = gpa, colour = at_risk_ind_term)) + 
    geom_point(data = toothInt, aes(y = val)) +
    geom_line(data = toothInt, aes(y = val, group = at_risk_ind_term)) + 
    apatheme 


#diagnose plots.
imputed_dom2_match_after.lm.res = imputed_dom2_match_after.lm
imputed_dom2_match_after.lm.res$M1.Fit = fitted(imputed_dom2_match_after.lm)
imputed_dom2_match_after.lm.res$M1.Resid = resid(imputed_dom2_match_after.lm)

ggplot(imputed_dom2_match_after.lm.res, aes(imputed_dom2_match_after.lm.res$M1.Fit, imputed_dom2_match_after.lm.res$M1.Resid, colour = wla_indicator)) + geom_point() +
xlab("Fitted Values") + ylab("Residuals") + apatheme

#-----------------------------------international students-----------------------------
imputed_intl2_match_after.lm <- lm(gpa ~ wla_indicator * at_risk_ind_term, data=imputed_intl2_match_after)
anova(imputed_intl2_match_after.lm)

#both the main effects and interaction effect are highly significant. 
coefficients(imputed_intl2_match_after.lm)
confint(imputed_intl2_match_after.lm)


apatheme=theme_bw()+
    theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        axis.line=element_line(),
        text=element_text(family='Times'))
#lsmeans.
leastsquare = lsmeans(imputed_intl2_match_after.lm, 
                      pairwise ~ wla_indicator:at_risk_ind_term,
                      adjust="tukey")
cld(leastsquare, 
    alpha=.05,  
    Letters=letters)

toothInt <- ddply(imputed_intl2_match_after,.(wla_indicator, at_risk_ind_term),summarise, val = mean(gpa))

#plots: main effect and interaction effect. 
#main effect.
ggplot(imputed_intl2_match_after, aes(x = at_risk_ind_term, y = gpa, color=wla_indicator)) + geom_boxplot() + scale_x_discrete() + xlab("at_risk_ind_term") + ylab("gpa") + apatheme 


ggplot(imputed_intl2_match_after.lm, aes(x = wla_indicator, y = gpa, colour = at_risk_ind_term)) + 
    geom_point(data = toothInt, aes(y = val)) +
    geom_line(data = toothInt, aes(y = val, group = at_risk_ind_term)) + 
    apatheme 


#diagnose plots.
imputed_intl2_match_after.lm.res = imputed_intl2_match_after.lm
imputed_intl2_match_after.lm.res$M1.Fit = fitted(imputed_intl2_match_after.lm)
imputed_intl2_match_after.lm.res$M1.Resid = resid(imputed_intl2_match_after.lm)

ggplot(imputed_intl2_match_after.lm.res, aes(imputed_intl2_match_after.lm.res$M1.Fit, imputed_intl2_match_after.lm.res$M1.Resid, colour = wla_indicator)) + geom_point() +
xlab("Fitted Values") + ylab("Residuals") + apatheme
```

###4.2 Dependent variable: graduation status. 
graduation_status = wla_indicator * at_risk_ind_term
Implication: wl improves the grauation probability by 2.3 times. At risk students seemed to be benefited the most by visitng wl. 
```{r echo=FALSE}
#-----------------------------------domestic students----------------------------------
temp <- subset(imputed_dom2_match_after, ttd_class_label %in% c('late', 'normal', 'dropped'))

#create a new indicator showing the graduation status.
temp$graduation_status <- ifelse(temp$ttd_class_label %in% c("late", "normal"), "1", "0")


temp$graduation_status <- as.factor(temp$graduation_status)

temp$graduation_status <- factor(temp$graduation_status,
levels = c(0,1),
labels = c("dropped", "graduated"))

temp$graduation_status <- as.factor(temp$graduation_status)
temp$graduation_status <- relevel(temp$graduation_status, ref = "dropped")


imputed_dom2_match_after.logistic2 <- glm(graduation_status ~ wla_indicator * at_risk_ind_term, family="binomial", data=temp)
summary(imputed_dom2_match_after.logistic2)

exp(coef(imputed_dom2_match_after.logistic2))

plot(allEffects(imputed_dom2_match_after.logistic2))

#-----------------------------------international students----------------------------------
temp <- subset(imputed_intl2_match_after, ttd_class_label %in% c('late', 'normal', 'dropped'))

#create a new indicator showing the graduation status.
temp$graduation_status <- ifelse(temp$ttd_class_label %in% c("late", "normal"), "1", "0")


temp$graduation_status <- as.factor(temp$graduation_status)

temp$graduation_status <- factor(temp$graduation_status,
levels = c(0,1),
labels = c("dropped", "graduated"))

temp$graduation_status <- as.factor(temp$graduation_status)
temp$graduation_status <- relevel(temp$graduation_status, ref = "dropped")


imputed_intl2_match_after.logistic2 <- glm(graduation_status ~ wla_indicator * at_risk_ind_term, family="binomial", data=temp)
summary(imputed_intl2_match_after.logistic2)

exp(coef(imputed_intl2_match_after.logistic2))

plot(allEffects(imputed_intl2_match_after.logistic2))
```

